Predicting Sakura Florescence in Japan with a Dual-Model Approach

### 2024-11-21 paper/Paper.qmd
Small changes to third draft of paper
Change all data source of tables, add train alternative model

# Install Packages
install.packages("tidyverse")
install.packages("arrow")
install.packages("here")
install.packages("ggplot2")
install.packages("patchwork")
install.packages("tidyverse")
install.packages("sf")
install.packages("rnaturalearth")
install.packages("rnaturalearthdata")
install.packages("ggthemes")
install.packages("dplyr")
install.packages("knitr")
install.packages("kableExtra")

# Load libraries
library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(patchwork)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggthemes)
library(dplyr)
library(knitr)
library(kableExtra)

@@

# Create a data frame for the coefficients and p-values
summary_table <- data.frame(
  Variable = c("latitude", "longitude"),
  Estimate = c(0.778101, -0.003319),
  P_Value = c("<2e-16", "0.806")
)

# Add R-squared information to the table
r_squared <- 0.6722
adjusted_r_squared <- 0.6721

# Combine R-squared information into the table
model_info <- data.frame(
  Variable = c("R-squared", "Adjusted R-squared"),
  Estimate = c(r_squared, adjusted_r_squared),
  P_Value = c("", "") # No p-value for R-squared metrics
)

# Combine both tables
final_table <- rbind(summary_table, model_info)

# Beautify the final_table using kable
kable(
  final_table,
  format = "latex",             # Output in LaTeX format
  booktabs = TRUE,              # Use booktabs style for clean table
  align = "c",                  # Center align columns
  col.names = c("Variable", "Estimate", "P-Value") # Column names
) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"), # Add striped rows and fixed position
    full_width = FALSE                             # Keep the table compact
  ) %>%
  add_header_above(c(" " = 1, "Model Summary" = 2)) # Add a header spanning the last two columns
```

@@

model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    "0.2337 on 572 degrees of freedom",
    "0.9987",
    "0.9987",
    "4.294e+05 on 1 and 572 DF",
    "<2e-16"
  )
)

# Model Summary Table for Model 1
kable(
  model_summary_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Metric", "Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Model Summary for Model 1" = 2))

@@

# Create a data frame for model summary
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    "1.116 on 4256 degrees of freedom",
    "0.9255",
    "0.9254",
    "8814 on 6 and 4256 DF",
    "<2e-16"
  )
)

# Model Summary Table for Model 2
kable(
  model_summary_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Metric", "Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Model Summary for Model 2" = 2))

@@

# Create a data frame for coefficients
coefficients_table <- data.frame(
  Variable = c("(Intercept)", "temp"),
  Estimate = c(126.202188, -3.509279),
  Std_Error = c(0.034396, 0.005355),
  T_value = c(3669.1, -655.3),
  P_Value = c("<2e-16", "<2e-16")
)

# Coefficients Table for Model 1
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Variable", "Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 1" = 5)) 
```

@@

# Create a data frame for coefficients
coefficients_table <- data.frame(
  Variable = c("(Intercept)", "day", "latitude", "month2", "month3", "month4", "month5"),
  Estimate = c(-9.000439, -0.049223, 0.189083, -0.617982, 2.869383, 6.966108, 10.047363),
  Std_Error = c(0.289464, 0.003291, 0.012089, 0.219831, 0.114364, 0.164984, 0.237453),
  T_value = c(-31.093, -14.956, 15.637, -2.811, 25.088, 42.223, 42.313),
  P_Value = c("<2e-16", "<2e-16", "<2e-16", "0.00496", "<2e-16", "<2e-16", "<2e-16")
)

# Coefficients Table for Model 2
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Variable", "Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 2" = 5)) 
```

@@

#### Beautify the comparison table using kable ####
comparison %>%
  kable(
    format = "latex",            # Output in LaTeX format
    booktabs = TRUE,             # Use booktabs style for clean table
    align = "c",                 # Center align columns
    col.names = c("Metric", "Actual", "Predicted")  # Column names
  ) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),  # Add striped rows and fixed position
    full_width = FALSE                              # Keep the table compact
  ) %>%
  add_header_above(c(" " = 1, "Prediction Summary" = 2))  # Add a header spanning last two columns

@@

# Create a data frame for residuals
residuals_table <- data.frame(
  Statistic = c("Min", "1Q", "Median", "3Q", "Max"),
  Value = c(-0.25208, -0.14651, -0.04095, 0.04606, 1.30709)
)

# Create a data frame for coefficients
coefficients_table <- data.frame(
  Variable = c("(Intercept)", "temp"),
  Estimate = c(126.202188, -3.509279),
  Std_Error = c(0.034396, 0.005355),
  T_value = c(3669.1, -655.3),
  P_Value = c("<2e-16", "<2e-16")
)

# Create a data frame for model summary
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    "0.2337 on 572 degrees of freedom",
    "0.9987",
    "0.9987",
    "4.294e+05 on 1 and 572 DF",
    "<2e-16"
  )
)

# Residuals Table for Model 1
kable(
  residuals_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c"
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Residuals for Model 1" = 2))

# Coefficients Table for Model 1
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Variable", "Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 1" = 5))

# Model Summary Table for Model 1
kable(
  model_summary_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Metric", "Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Model Summary for Model 1" = 2))

@@

# Create a data frame for residuals
residuals_table <- data.frame(
  Statistic = c("Min", "1Q", "Median", "3Q", "Max"),
  Value = c(-4.2185, -0.6993, -0.0729, 0.6641, 4.1114)
)

# Create a data frame for coefficients
coefficients_table <- data.frame(
  Variable = c("(Intercept)", "day", "latitude", "month2", "month3", "month4", "month5"),
  Estimate = c(-9.000439, -0.049223, 0.189083, -0.617982, 2.869383, 6.966108, 10.047363),
  Std_Error = c(0.289464, 0.003291, 0.012089, 0.219831, 0.114364, 0.164984, 0.237453),
  T_value = c(-31.093, -14.956, 15.637, -2.811, 25.088, 42.223, 42.313),
  P_Value = c("<2e-16", "<2e-16", "<2e-16", "0.00496", "<2e-16", "<2e-16", "<2e-16")
)

# Create a data frame for model summary
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    "1.116 on 4256 degrees of freedom",
    "0.9255",
    "0.9254",
    "8814 on 6 and 4256 DF",
    "<2e-16"
  )
)

@@

# Coefficients Table
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Variable", "Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 2" = 5))

>>>>

# Install Packages
install.packages("tidyverse")
install.packages("arrow")
install.packages("here")
install.packages("ggplot2")
install.packages("patchwork")
install.packages("tidyverse")
install.packages("sf")
install.packages("rnaturalearth")
install.packages("rnaturalearthdata")
install.packages("ggthemes")
install.packages("dplyr")
install.packages("knitr")
install.packages("kableExtra")
# Load libraries
library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(patchwork)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggthemes)
library(dplyr)
library(knitr)
library(kableExtra)

@@

# Load necessary library
library(knitr)
library(kableExtra)

# Train the model
alternative_model <- lm(mean_temp_month ~ latitude + longitude, data = sakura_modern_cleaned)

# Extract coefficients and p-values
coefficients_summary <- summary(alternative_model)$coefficients
coefficients_table <- data.frame(
  Variable = rownames(coefficients_summary),
  Estimate = coefficients_summary[, "Estimate"],
  P_Value = format.pval(coefficients_summary[, "Pr(>|t|)"], digits = 4) # Properly formatted p-values
)

# Extract R-squared and Adjusted R-squared
r_squared <- summary(alternative_model)$r.squared
adjusted_r_squared <- summary(alternative_model)$adj.r.squared

# Add R-squared information to the table
model_info <- data.frame(
  Variable = c("R-squared", "Adjusted R-squared"),
  Estimate = c(r_squared, adjusted_r_squared),
  P_Value = c("", "") # No p-value for R-squared metrics
)

# Combine coefficients and model information
final_table <- rbind(coefficients_table, model_info)

# Generate LaTeX table
kable(
  final_table,
  format = "latex",             # Output in LaTeX format
  booktabs = TRUE,              # Use booktabs style for clean table
  align = "c",                  # Center align columns
  col.names = c("Variable", "Estimate", "P-Value") # Column names
) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"), # Add striped rows and fixed position
    full_width = FALSE                             # Keep the table compact
  )

@@

# Calculate p-value for the F-statistic dynamically
f_stat_1 <- summary(model1)$fstatistic
p_value_1 <- pf(f_stat_1[1], f_stat_1[2], f_stat_1[3], lower.tail = FALSE)

# Create a data frame for model summary (dynamically extracted values)
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    paste0(round(summary(model1)$sigma, 4), " on ", summary(model1)$df[2], " degrees of freedom"), # Residual Standard Error
    round(summary(model1)$r.squared, 4), # Multiple R-squared
    round(summary(model1)$adj.r.squared, 4), # Adjusted R-squared
    paste0(round(f_stat_1[1], 4), " on ", f_stat_1[2], " and ", f_stat_1[3], " DF"), # F-statistic
    format.pval(p_value_1, digits = 4) # Properly formatted p-value
  )
)

# Generate LaTeX table for Model 1 summary
kable(
  model_summary_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Metric", "Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

@@

# Calculate p-value for the F-statistic dynamically
f_stat_2 <- summary(model2)$fstatistic
p_value_2 <- pf(f_stat_2[1], f_stat_2[2], f_stat_2[3], lower.tail = FALSE)

# Create a data frame for model summary (dynamically extracted values)
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    paste0(round(summary(model2)$sigma, 4), " on ", summary(model2)$df[2], " degrees of freedom"), # Residual Standard Error
    round(summary(model2)$r.squared, 4), # Multiple R-squared
    round(summary(model2)$adj.r.squared, 4), # Adjusted R-squared
    paste0(round(f_stat_2[1], 4), " on ", f_stat_2[2], " and ", f_stat_2[3], " DF"), # F-statistic
    format.pval(p_value_2, digits = 4) # Properly formatted p-value
  )
)

# Generate LaTeX table for Model 2 summary
kable(
  model_summary_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Metric", "Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

@@

# Extract coefficients summary and format p-values properly
coefficients_summary <- summary(model1)$coefficients
coefficients_table <- data.frame(
  Estimate = coefficients_summary[, "Estimate"], # Extract estimates
  Std_Error = coefficients_summary[, "Std. Error"], # Extract standard errors
  T_value = coefficients_summary[, "t value"], # Extract t-values
  P_Value = format.pval(coefficients_summary[, "Pr(>|t|)"], digits = 4) # Proper p-value formatting
)

# Coefficients Table for Model 1
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

@@

# Extract coefficients summary and format p-values properly
coefficients_summary <- summary(model2)$coefficients
coefficients_table <- data.frame(
  Estimate = coefficients_summary[, "Estimate"], # Extract estimates
  Std_Error = coefficients_summary[, "Std. Error"], # Extract standard errors
  T_value = coefficients_summary[, "t value"], # Extract t-values
  P_Value = format.pval(coefficients_summary[, "Pr(>|t|)"], digits = 4) # Proper p-value formatting
)

# Coefficients Table for Model 2
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

@@

#### Beautify the comparison table using kable ####
comparison %>%
  kable(
    format = "latex",            # Output in LaTeX format
    booktabs = TRUE,             # Use booktabs style for clean table
    align = "c",                 # Center align columns
    col.names = c("Metric", "Actual", "Predicted")  # Column names
  ) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),  # Add striped rows and fixed position
    full_width = FALSE                              # Keep the table compact
  )

@@

# Extract residuals statistics
residuals_table <- data.frame(
  Statistic = c("Min", "1Q", "Median", "3Q", "Max"),
  Value = quantile(residuals(model1), probs = c(0, 0.25, 0.5, 0.75, 1))
)

# Extract coefficients summary and format p-values properly
coefficients_summary <- summary(model1)$coefficients
coefficients_table <- data.frame(
  Estimate = coefficients_summary[, "Estimate"],
  Std_Error = coefficients_summary[, "Std. Error"],
  T_value = coefficients_summary[, "t value"],
  P_Value = format.pval(coefficients_summary[, "Pr(>|t|)"], digits = 4) # Proper p-value formatting
)

# Calculate p-value for the F-statistic dynamically
f_stat_1 <- summary(model1)$fstatistic
p_value_1 <- pf(f_stat_1[1], f_stat_1[2], f_stat_1[3], lower.tail = FALSE)

# Extract model summary metrics
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    paste0(round(summary(model1)$sigma, 4), " on ", summary(model1)$df[2], " degrees of freedom"),
    round(summary(model1)$r.squared, 4),
    round(summary(model1)$adj.r.squared, 4),
    paste0(round(f_stat_1[1], 4), " on ", f_stat_1[2], " and ", f_stat_1[3], " DF"),
    format.pval(p_value_1, digits = 4) # Format p-value dynamically
  )
)

# LaTeX table generation remains unchanged
kable(
  residuals_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c"
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Residuals for Model 1" = 2))

kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 1" = 5))

kable(
  model_summary_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Metric", "Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Model Summary for Model 1" = 2))

@@

# Extract residuals statistics from Model 2
residuals_table <- data.frame(
  Statistic = c("Min", "1Q", "Median", "3Q", "Max"),
  Value = quantile(residuals(model2), probs = c(0, 0.25, 0.5, 0.75, 1))
)

# Extract coefficients summary and format p-values properly
coefficients_summary <- summary(model2)$coefficients
coefficients_table <- data.frame(
  Estimate = coefficients_summary[, "Estimate"],
  Std_Error = coefficients_summary[, "Std. Error"],
  T_value = coefficients_summary[, "t value"],
  P_Value = format.pval(coefficients_summary[, "Pr(>|t|)"], digits = 4) # Proper p-value formatting
)

# Calculate p-value for the F-statistic dynamically
f_stat_2 <- summary(model2)$fstatistic
p_value_2 <- pf(f_stat_2[1], f_stat_2[2], f_stat_2[3], lower.tail = FALSE)

# Extract model summary metrics
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    paste0(round(summary(model2)$sigma, 4), " on ", summary(model2)$df[2], " degrees of freedom"),
    round(summary(model2)$r.squared, 4),
    round(summary(model2)$adj.r.squared, 4),
    paste0(round(f_stat_2[1], 4), " on ", f_stat_2[2], " and ", f_stat_2[3], " DF"),
    format.pval(p_value_2, digits = 4) # Format p-value dynamically
  )
)

# LaTeX table generation remains unchanged

@@

# Coefficients Table
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 2" = 5))
